waRne - Putting cricket analytics in a spin

### 2020-09-26 paper.Rmd
keep on writing

---
title: "waRne - Putting cricket analytics in a spin"
authors: Robert(1), James (2)
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true

@@

## What is waRne

We need to think about what it is waRne should be?

1.  provide the csv we used for datajam and then people just add games to it

2.  people can not have the csv but just download random tournaments

@@

@@

James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season. The statistic given was in the 2019 big bash season teams that batted first won only $43$%\$ of games. Thankfully we can try to reproduce that statistic now through the use of waRne

@@

@@

: things to do

## How would you recommend evaluating the selection of players to reduce the required run rate? Is there a tool that would could be used to let coaches or selectors do this themselves or is it would it be more appropriate to have an analyst do this prior?

##

>>>>

---
title: "waRne - Putting cricket analytics in a spin"
authors: Robert(1), James (2)
output: 
  html_document:
    toc: true
    toc_float: true
    keep_md: true
bibliography: references.bib

@@

Data access is a key enabler for any analytics community. Most major sports have easy access to match statistics, for example [nbastatR](https://github.com/abresler/nbastatR) for NBA, [Lahman](https://cran.r-project.org/web/packages/Lahman/Lahman.pdf) for baseball, [deuce](https://github.com/skoval/deuce) for tennis and [nflfastR](https://github.com/mrcaseb/nflfastR) for NFL. Through access to sports data and reproducible findings and metrics fans clubs and researchers are better able to understand the game, predict match outcomes and rate players. For example the way teams tackled 4th down decisions in the NFL has changed since [@romer2006firms] seminial work. As teams have changed their 4th down decision making this allows follow up research such as [@yam2019lost] which looked at more granular data to see if this happens in practice.

By making data more accessible and more advanced metrics more accessible fans data journalism in sports has grown in recent years. For example in [@horowitz2017nflscrapr] has enabled EPA to enter popular discussion among fans.

Cricket is the second largest sport in the world. However, unfortunetaly there is no easy accessible way to access ball by ball data nor aggregated statistics of teams and players. Data while available on sites like [espncricinfo](https://www.espncricinfo.com/) are not in an easy to use form. For example, each match is listed on different webpages so hours upon hours of time would be required to copy and paste a single season. Hence, there are significant logistical barriers for prospective fans and analysts studying the game, which stagnats understanding of cricket.

This paper describes the waRne package, the first to provide free and easy access to data for cricket for fans. Web scraping tools are avaiable for fans to easily scrape the play by play commentary data on espncricinfo. For the first time fans can evaluate their favourite teams and players and do so in a reproducible and accessible manner.

@@

Data accessibility enables fans, analyts and researchers to better understand the game. Through being able to reproduce common popular metrics, visualisations and article findings.

Through being able to reproduce, fans are able to make accessible findings for others and importantly they are able to extend and grow concepts. Unfortuntely what we see through leading cricket analytics providers is a track record of confusing output for fans. This can lead to lower engagement and dismissial of cricket analytics.

@@

James and I entered a reproducibility competition and the winning entry provided a statistic sort of a gateway into the importance of their ''project'' was was specifically in 2019 Australian Big Bash T20 season. The statistic given was in the 2019 big bash season teams that batted first won only $43$% of games. Thankfully we can try to reproduce that statistic now through the use of waRne

@@

## Easier engagement of fans

To the surprise of many, its hard to engage cricket fans into the analytics behind the game. Reasons for this are generally centred around reproducibility and explainability to fans.

Without an easy accessible medium how can crickets version of an analytics community grow.

something something look towards how EPA has changed the way fans are engaged in NFL analytics - so maybe the change in run rate can be like EPA?

```{r}
library(tidyverse)
df<-read.csv("data/combined_2020_09_11.csv")

df%>%
	group_by(match_id)%>%
	filter(inning==2)%>%
	group_by()

@@

## Dhoni Dilemma

With this new dataset, we allow fans to put on their analyst thats and to be able to ask themselves, what exactly makes a good closer and what exactly is the Dhoni Dilemma.
